---
title: "Cocktail bair"
author: "Chauke Helga"
date: "2022-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(dplyr)
```

## 1. Read the data
```{r}
url <- "https://raw.githubusercontent.com/nthun/cocktail-balance/master/cocktail_data.tsv"
ctbar <- readr::read_tsv(url)
```
## 2. Transform the data table and clean the ingredient variable!
```{r}
fnl<- ctbar %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = "[1234567890,.</()%]",replacement = ""))

fnl2<- fnl %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = "drops", replacement = ""))
fnl3<- fnl2 %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = "drop", replacement = ""))
fnl4<- fnl3 %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = "dashes", replacement = ""))
fnl5<- fnl4 %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = "abvbr", replacement = ""))
fnl6<- fnl5 %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = "br", replacement = ""))
fnl7<- fnl6 %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = "abv", replacement = ""))
fnl8<- fnl7 %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = "oz", replacement = ""))
fnl9<- fnl8 %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = "dash", replacement = ""))
fnl0<- fnl9 %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = "bsp", replacement = ""))


```

```{r}
 df1 <- fnl0 %>% tidyr::separate_rows(ingredients, sep = ">")
df1a<- df1 %>% mutate(ingredients = str_trim(df1$ingredients, "both"))

#trm <- gsub("^\\s+|\\s+$", "", df1$ingredients)
#df1c<- df1 %>% mutate(ingredients = str_replace_all(string = ingredients, pattern = " ", replacement = ""))



```
```{r}

```

## 3. All ingredients in alphabetical order
```{r}
sort1 <- with(df1a,  df1a[order(ingredients) , ])
```

## 4. Number of unique ingredients
```{r}
n_distinct(sort1$ingredients)

count <- sort1 %>%group_by(ingredients) %>% summarise(count = n_distinct(ingredients))
```
## 5. What are the top 10 ingredients?
```{r}
names(which.max(table(df1a$ingredients)))

lvls <- unique(unlist(df1a$ingredients))
 freq <- sapply(df1a, function(x) table(factor(x, levels = lvls,ordered = TRUE)))
 print ("Count of variables per column")
 print (freq)
 
#ingcount <- order((freq$ingredients) , ])

 #freq = lemon juice, simple syrup, sweet vermouth, gin, lime juice, Angostura bitters, saline solution, water, rye, cognac 

```
## 6. Which cocktail(s) has/have the most ingredients?
```{r}
lvls <- unique(unlist(df1a$name))
namefreq <- sapply(df1a, function(x) table(factor(x, levels = lvls,ordered = TRUE)))
#mostingredients <-pink lady, clover cubm viex carre, blender margarita, carbonated negroni

table(df1a$name)
```


## 7. How many ingredients appear in only one cocktail (rare ingredient)?
```{r}
names(which.min(table(df1a$name)))
```
## 8. Which cocktail has an ingredient that is only used in one cocktail?
```{r}
d2 <- df1a %>% group_by(ingredients) %>% dplyr::filter(n() == 1) %>%ungroup()
```
## 9. What are the cocktails without rare ingredients?
```{r}
print("Elements which are repeated")
which(table(df1a$name)>1)
```
## 10. Create a cheat sheet for the bartender!
```{r}
mtr <-as_tibble(ctbar$name)
mtr2 <- as_tibble(count$ingredients)
mtr3 <- t(mtr2)
colnames(mtr3) <- mtr3[1,]
cs <- cbind(mtr,mtr3)

cs[,]=matrix(ncol= ncol(cs), rep("", prod(dim(cs))))
cs1 <- cbind(mtr,cs)
cs1a <- cs1[,-2]

```

