---
title: "Indoor Air Pollution"
author: "Chauke Helga"
date: "2022-12-21"
output:
  html_document: default
  pdf_document: default
---

```{r, error=TRUE}

knitr::opts_chunk$set(echo = TRUE)
```

###packages
```{r}
library(explore)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(rpart)
library(forecast)
```

#read data
#remame column 4
```{r}
indoor_pollution <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-04-12/indoor_pollution.csv')


```
##explore the data
```{r}
indoor_pollution %>% explore_tbl()
indoor_pollution %>% describe()
summary(indoor_pollution[, "Entity"])

```
##rename col4
```{r}
colnames(indoor_pollution)[4] <- "Deaths_pcnt"
```

#calculate world statistics
```{r}
mean(indoor_pollution$Deaths_pcnt)
median(indoor_pollution$Deaths_pcnt)
Worldstats <- na.exclude(indoor_pollution$Deaths_pcnt)
sum(indoor_pollution$Deaths_pcnt)
var(indoor_pollution$Deaths_pcnt, na.rm=TRUE)
sd(indoor_pollution$Deaths_pcnt, na.rm = TRUE)
coefVar <- sd(indoor_pollution$Deaths_pcnt) / mean(indoor_pollution$Deaths_pcnt) * 100
quantile(indoor_pollution$Deaths_pcnt)
range(indoor_pollution$Deaths_pcnt)
IQR(indoor_pollution$Deaths_pcnt)
```
#world data histogram and density plot
```{r}
# histogram
ggplot(indoor_pollution, aes(x = Deaths_pcnt)) +
  geom_histogram(bins = nclass.Sturges(indoor_pollution$Deaths_pcnt))

# density curve
ggplot(indoor_pollution, aes(x = Deaths_pcnt)) + geom_density()

indoor_pollution%>% select(Entity, Year, Deaths_pcnt ) %>% 
  explore_all(target = Entity)

ggplot(indoor_pollution, aes(x = Year, y = Deaths_pcnt, color = Entity, group = Entity)) +
  geom_line() + geom_point() +   theme_test()
```
##world total
```{r}
Worldtot <- indoor_pollution %>% group_by(Year) %>% summarise(Deaths_pcnt = sum(Deaths_pcnt))



ggplot(Worldtot, aes(x = Year, y= Deaths_pcnt)) + geom_line()
ggplot(Worldtot, aes(x = Deaths_pcnt)) + geom_histogram(bins = nclass.Sturges(Worldtot$Deaths_pcnt))

 ggplot(indoor2, aes(x= Year, Deaths_pcnt, color = Entity)) + geom_line()+ theme(legend.position = "none")
```
##number of deaths around the world have decreased with time, but which country still experience the highest number of deaths in moden times?
```{r}

 lastyear <- filter(indoor_pollution, Year ==2019)

sapply(lastyear, max) 
 
 ggplot(lastyear, aes(x= Deaths_pcnt, y= Entity)) + geom_col()
 
 
indoor2 <-select(indoor_pollution, -c(Code))
worlddeath <- spread(indoor2, key = Entity, value = Deaths_pcnt)

sapply(lastyear, min)

```

#compute statistical model for world data
##is there significant difference between the first year vs the last year of data collection in the world?
```{r}
 worlddeathYr <- spread(indoor2, key = Year, value = Deaths_pcnt)

t.test(worlddeathYr$`2019`, worlddeathYr$`1990`, var.equal = F)

#ttest of the last year
t.test(lastyear$Deaths_pcnt,mu=0)


###ANOva

res_aov <- aov(Deaths_pcnt ~ Entity,
               data = indoor2)

  hist(res_aov$residuals)
  
boxplot()
```


###residual diagnostics
```{r}
 hist(res_aov$residuals)
    
        checkresiduals(indoor2$Deaths_pcnt)
        checkresiduals(res_aov$residuals)
```

